### 2023-08-31 sections/1-introduction.tex
getting there

In particular, $k$-nearest neighbors ($k$-NN) search is one of the most pervasive classification and recommendation methods in use~\cite{fix1952discriminatory, cover1967nearest}; 
Naive implementations of kNN, whose time complexity is linear in the dataset's cardinality, 
prove prohibitively slow for large datasets. While algorithms for fast $k$-NN on large datasets exist, most of them are approximate and do not 
exploit the geometric and topological structure inherent in big datasets. Approximate search is sufficient for some applications, but the 
need for efficient, \emph{exact} search remains.
In particular, while approximate $k$-NN may agree with exact $k$-NN in a majority vote for large values of $k$, it may be sensitive to local perturbations with smaller values of $k$.

@@

This paper focuses on accelerated $\rho$- and $k$-nearest neighbors search. We explain some 
implementation improvements to CHESS, as well as improved genericity across distance functions. 
We also include proof of some geometric guarantees for clusters. Finally, we provide a comparison of 
CAKES to the state-of-the-art in similarity search, FAISS~\cite{johnson2019billion} on several datasets 
from the ann and big ann benchmarks suites.

>>>>

In particular, $k$-nearest neighbors ($k$-NN) search is one of the most pervasive classification and recommendation methods in use~\cite{fix1952discriminatory, cover1967nearest}; 
naive implementations of kNN, whose time complexity is linear in the dataset's cardinality, 
prove prohibitively slow for large datasets. While algorithms for fast $k$-NN on large datasets exist, most of them are approximate and do not 
exploit the geometric and topological structure inherent in big datasets. Approximate search is sufficient for some applications, but the 
need for efficient, \emph{exact} search remains.
In particular, while approximate $k$-NN may agree with exact $k$-NN in a majority vote for large values of $k$, it may be sensitive to local perturbations with smaller values of $k$.

@@

This paper focuses on four novel accelerated $k$-nearest neighbors search algorithms. We explain some 
implementation improvements to CHESS, as well as improved genericity across distance functions. 
We also include proof of some geometric guarantees for clusters. Finally, we provide a comparison of 
CAKES to the state-of-the-art in similarity search, FAISS~\cite{johnson2019billion} on several datasets 
from the ann and big ann benchmarks suites.
